---
title: "Case II only Samsung"
output: html_document
date: "2025-04-08"
---

```{r, include=FALSE}

# Сlear the database
rm(list=ls()) 
graphics.off()

```

```{r setup, include=FALSE, results = "hide"}

#Readability in HTML report
#install.packages("DT")
library(DT)

# Clear the database
rm(list=ls()) 

#Install necessary packages. Delete # sign in order to download all of them. 
#install.packages('rcompanion', repos = "https://cloud.r-project.org")
#install.packages('nortest', repos = "https://cloud.r-project.org")
#install.packages('corrplot', repos = "https://cloud.r-project.org")
#install.packages('olsrr', repos = "https://cloud.r-project.org")
#install.packages('dplyr', repos = "https://cloud.r-project.org")
#install.packages('pastecs', repos = "https://cloud.r-project.org")
#install.packages('REdaS', repos = "https://cloud.r-project.org")
#install.packages('psych', repos = "https://cloud.r-project.org")
#install.packages('lm.beta', repos = "https://cloud.r-project.org")


# Load required libraries
library(ggplot2)
library(dplyr)
library(readr)
library(broom)
library(lm.beta)
library(pastecs)
library(tidyr)
library(corrplot)     # For correlation matrix visualization
library(olsrr)        # For VIF and tolerance values
library(REdaS)        # For Bartlett's test
library(psych)        # For PCA and KMO

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Read the dataset directly from the provided URL
url <- "https://raw.githubusercontent.com/silinette/Case-Study-II/refs/heads/main/Data%20File_Case_Study_Factor%20Analysis_MD.csv"

Data <- read.csv(url, header = TRUE)
Data
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Explore the structure of the dataset
dim(Data) # Check the number of rows and columns in the dataset
head(Data) # Display the first few rows of the dataset
str(Data) # Display the structure of the dataset (types, sample values)
names(Data) # Show all variable names in the dataset
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Check for missing values
list_na <- colnames(Data)[apply(Data, 2, anyNA)] # Identify columns with missing values
list_na 
sum(is.na(Data)) # Count total number of missing values
sum(!is.na(Data)) # Count total number of non-missing values

#Column names that we do not touch/Define variables we will keep even with NAs:
cols_keep_na <- c("noOfDefects", "defectsWarranty", "fRecall",
                  "def_sev1", "def_sev2", "didComplain",
                  "reactionComplain", "nextPurchase")

# Create column vector to check for clearances/Define variables to check for complete cases:
cols_check <- setdiff(names(Data), cols_keep_na)

# Filter for Samsung only (brandrec == 2) and apply listwise deletion
Data_clean <- Data %>%
  filter(brandrec == 2) %>%                # keep only Samsung
  filter(complete.cases(across(all_of(cols_check))))  # listwise deletion


#write.csv(Data_clean, file = "~/Desktop/Data_clean.csv", row.names = FALSE)

```
### Conclusion 

The results of the factor analysis revealed nine distinct dimensions of smartphone quality as perceived by consumers: Performance, Features, Reliability, Durability, Ease of Use, Service, Aesthetics, Distinctiveness, and Conformance. Each factor grouped conceptually similar survey items and aligned well with the existing literature.

The scree plot and eigenvalues supported the retention of 9 factors, cumulatively explaining over 82% of the total variance. Varimax rotation helped in achieving clearer interpretation by minimizing cross-loadings.

Using factor scores, we compared average brand perceptions across these dimensions. The differences observed can be used to guide strategic decisions regarding product improvement and marketing communication.

This analysis demonstrates how factor analysis can uncover meaningful insights about consumer perceptions that go beyond simple satisfaction metrics. It provides a powerful tool for product benchmarking and positioning in a competitive market.
